\chapter{API Endpoint Reference}
\label{appx:api}

The Seed Bank backend exposes a single versioned REST surface mounted under
\texttt{/api/v1} \cite{fielding2000rest}. Routers are split by resource, and the
interactive OpenAPI documentation is served at \texttt{/api/v1/docs} with the raw
schema at \texttt{/api/v1/openapi.json}; the web client's typed request layer is
generated from that schema, so the table below and the running code cannot drift
apart. Every successful single response is wrapped in a \texttt{\{ "data": ... \}}
envelope, list responses add a \texttt{meta} block with pagination, and every error
is an RFC 9457 problem-details document carrying the request's \texttt{request\_id}
\cite{rfc9457problem}.

Table~\ref{tab:appx-api} lists the representative endpoints grouped by router.
Paths are shown relative to the \texttt{/api/v1} prefix. Authentication uses a
short-lived JWT access token \cite{rfc7519jwt}, except for the public
\texttt{shared} route and the OAuth handshake \cite{rfc6749oauth}.

\begin{singlespace}
\begin{longtable}{@{}l l p{6.4cm}@{}}
\caption{Representative REST endpoints by router (paths relative to
\texttt{/api/v1}).}\label{tab:appx-api}\\
\toprule
\textbf{Method} & \textbf{Path} & \textbf{Purpose} \\
\midrule
\endfirsthead
\multicolumn{3}{@{}l}{\small\itshape Table~\ref{tab:appx-api} continued from previous page}\\
\toprule
\textbf{Method} & \textbf{Path} & \textbf{Purpose} \\
\midrule
\endhead
\midrule
\multicolumn{3}{r@{}}{\small\itshape continued on next page}\\
\endfoot
\bottomrule
\endlastfoot

\multicolumn{3}{@{}l}{\textbf{auth}}\\
POST   & \texttt{/auth/register}        & Create a local account; sends a verification email. \\
POST   & \texttt{/auth/verify-email}    & Confirm an account from an emailed token. \\
POST   & \texttt{/auth/login}           & Exchange email and password for access + refresh tokens. \\
POST   & \texttt{/auth/refresh}         & Rotate the refresh token and mint a new access token. \\
POST   & \texttt{/auth/logout}          & Revoke the current refresh-token chain. \\
GET    & \texttt{/auth/oauth/\{provider\}/login}    & Start a Google or GitHub OAuth flow. \\
GET    & \texttt{/auth/oauth/\{provider\}/callback} & Complete the OAuth redirect and issue tokens. \\
POST   & \texttt{/auth/bootstrap-admin} & One-shot first-admin creation, guarded by a shared token. \\
\addlinespace
\multicolumn{3}{@{}l}{\textbf{users}}\\
GET    & \texttt{/users/me}             & Return the authenticated user's profile. \\
GET    & \texttt{/users}                & Admin: list users with pagination. \\
PATCH  & \texttt{/users/\{user\_id\}/role} & Admin: change a user's role. \\
\addlinespace
\multicolumn{3}{@{}l}{\textbf{api\_keys}}\\
POST   & \texttt{/api-keys}             & Create a personal key (returned in plaintext once). \\
GET    & \texttt{/api-keys}             & List the caller's keys (metadata only). \\
DELETE & \texttt{/api-keys/\{key\_id\}}  & Revoke a key. \\
\addlinespace
\multicolumn{3}{@{}l}{\textbf{models}}\\
POST   & \texttt{/models}                       & Register a model artifact and its weights/metadata. \\
GET    & \texttt{/models}                       & List registered models. \\
GET    & \texttt{/models/\{model\_id\}}         & Fetch metadata for a specific model. \\
PATCH  & \texttt{/models/\{model\_id\}}         & Advance lifecycle status (staging/production/archived). \\
GET    & \texttt{/models/\{model\_id\}/performance} & Per-model offline metrics and ClickHouse analytics. \\
\addlinespace
\multicolumn{3}{@{}l}{\textbf{traffic}}\\
GET    & \texttt{/traffic-splits}       & Admin: list active splits. \\
PATCH  & \texttt{/traffic-splits}       & Admin: atomically replace splits for a segment. \\
\addlinespace
\multicolumn{3}{@{}l}{\textbf{analyze}}\\
POST   & \texttt{/analyze}              & Multipart image upload that starts an inference batch. \\
\addlinespace
\multicolumn{3}{@{}l}{\textbf{batches}}\\
GET    & \texttt{/batches}                       & List the caller's batches. \\
GET    & \texttt{/batches/\{batch\_id\}}          & Poll one batch for status and results. \\
GET    & \texttt{/batches/\{batch\_id\}/image-urls} & Fetch short-lived presigned URLs for batch's images. \\
POST   & \texttt{/batches/compare}               & Compare aggregate stats side-by-side for 2--10 batches. \\
DELETE & \texttt{/batches/\{batch\_id\}}          & Soft-delete one batch. \\
POST   & \texttt{/batches/delete}                & Soft-delete up to 200 owned batches in bulk. \\
GET    & \texttt{/batches/\{batch\_id\}/export.csv}  & Export all detections in the batch as CSV. \\
GET    & \texttt{/batches/\{batch\_id\}/export.json} & Export all detections in the batch as JSON. \\
GET    & \texttt{/batches/\{batch\_id\}/images/\{image\_id\}/annotated.png} & Get original image with detection boxes burned in. \\
POST   & \texttt{/batches/\{batch\_id\}/share}    & Create or rotate public read-only share link. \\
DELETE & \texttt{/batches/\{batch\_id\}/share}    & Revoke the share link. \\
\addlinespace
\multicolumn{3}{@{}l}{\textbf{analytics}}\\
GET    & \texttt{/analytics}            & Aggregated metrics over a time window (ClickHouse). \\
\addlinespace
\multicolumn{3}{@{}l}{\textbf{shared}}\\
GET    & \texttt{/shared/\{token\}}     & Public, unauthenticated read-only batch report. \\
\addlinespace
\multicolumn{3}{@{}l}{\textbf{catalog}}\\
GET    & \texttt{/seed-types}           & Reference list of seed types. \\
GET    & \texttt{/suppliers}            & Reference list of visible suppliers. \\
POST   & \texttt{/suppliers}            & Create a global (admin) or private supplier. \\
PATCH  & \texttt{/suppliers/\{supplier\_id\}} & Update a supplier's metadata or active status. \\
DELETE & \texttt{/suppliers/\{supplier\_id\}} & Soft-delete a supplier. \\
\addlinespace
\multicolumn{3}{@{}l}{\textbf{datasets}}\\
POST   & \texttt{/datasets}                 & Create a labelled dataset for offline evaluation. \\
GET    & \texttt{/datasets}                 & List datasets. \\
GET    & \texttt{/datasets/\{dataset\_id\}} & Fetch metadata for a specific dataset. \\
POST   & \texttt{/datasets/\{dataset\_id\}/items} & Add ground-truth items to a dataset. \\
GET    & \texttt{/datasets/\{dataset\_id\}/items} & List items in a dataset with pagination. \\
\addlinespace
\multicolumn{3}{@{}l}{\textbf{experiments}}\\
POST   & \texttt{/experiments}              & Create and run an offline-evaluation experiment. \\
GET    & \texttt{/experiments}              & List experiments. \\
GET    & \texttt{/experiments/\{experiment\_id\}} & Fetch experiment status and summary. \\

\end{longtable}
\end{singlespace}


\chapter{Requirements Traceability}
\label{appx:req}

The consolidated functional and non-functional requirements are defined once, in
Chapter~3. The use-case diagram appears as \autoref{fig:usecase}, and the
functional (FR) and non-functional (NFR) requirement identifiers are listed in the
FR and NFR tables of that chapter. To avoid two copies that can disagree, they are
not reprinted here.

For grading and review, the mapping from requirement to implementation is direct.
Each FR resolves to one router in Appendix~A: the analyze and batch requirements
map to the \texttt{analyze} and \texttt{batches} routers, the model-management and
A/B requirements map to the \texttt{models} and \texttt{traffic} routers, and the
identity requirements map to \texttt{auth}, \texttt{users}, and \texttt{api\_keys}.
The data-traceability requirement maps to the four-table inference chain
(\texttt{scan\_batches} $\rightarrow$ \texttt{scan\_images} $\rightarrow$ 
\texttt{inferences} $\rightarrow$ \texttt{seed\_detections}) described in
Chapter~4. The non-functional requirements map to mechanisms documented elsewhere
in the report: concurrency safety to the compare-and-set batch state machine
(Appendix~D), observability to the Prometheus, OpenTelemetry, and Sentry signals,
and reproducibility on the synthetic-data side to the seeded MultiSeedGen pipeline
(Appendix~C).


\chapter{MultiSeedGen Configuration Reference}
\label{appx:msg}

MultiSeedGen is the companion tool that turns single-seed source photos into
multi-seed detection datasets by cut-and-paste compositing
\cite{dwibedi2017cutpaste}. It is driven the same way from a CLI flag, a
YAML/JSON config file, or the web form, because all three feed one validated
\texttt{Config}; CLI flags override file values, and every run writes a
\texttt{run\_config.yaml} that reproduces the dataset when fed back in. The
recommended workflow is to verify segmentation first, generate, then validate the
exported labels.

Table~\ref{tab:appx-msg} lists the knobs that matter most in practice, with their
defaults. Quality of the cut-outs and the domain match of the backgrounds
dominate dataset quality far more than image count, so the segmentation and
background flags are the ones worth tuning first.

\begin{singlespace}
\begin{longtable}{@{}l l p{6.6cm}@{}}
\caption{Most useful MultiSeedGen configuration flags and their effect.}
\label{tab:appx-msg}\\
\toprule
\textbf{Flag} & \textbf{Default} & \textbf{Effect} \\
\midrule
\endfirsthead
\multicolumn{3}{@{}l}{\small\itshape Table~\ref{tab:appx-msg} continued from previous page}\\
\toprule
\textbf{Flag} & \textbf{Default} & \textbf{Effect} \\
\midrule
\endhead
\midrule
\multicolumn{3}{r@{}}{\small\itshape continued on next page}\\
\endfoot
\bottomrule
\endlastfoot

\texttt{-{}-segment}            & \texttt{auto}   & Cut-out method: \texttt{auto} (classical cascade with a learned fallback), \texttt{rembg} (U2-Net matting \cite{qin2020u2net}), \texttt{grabcut}, \texttt{otsu}, \texttt{u2net}. \\
\texttt{-{}-seg-conf-thresh}    & \texttt{0.55} / \texttt{0} & Skip-gate for low-confidence cut-outs. Active in \texttt{auto}; opt-in (defaults to \texttt{0}, off) for explicit methods. \\
\texttt{-{}-rembg-provider}     & \texttt{auto}   & Run the learned matting on GPU when present (about $11\times$ faster), CPU otherwise. \\
\texttt{-{}-min-seeds}          & n/a             & Lower bound on seeds composited per scene. \\
\texttt{-{}-max-seeds}          & n/a             & Upper bound on seeds per scene (must be $\geq$ \texttt{min-seeds}). \\
\texttt{-{}-scale-min}          & n/a             & Smallest per-seed scale as a fraction of the canvas. \\
\texttt{-{}-scale-max}          & n/a             & Largest per-seed scale (controls apparent seed size). \\
\texttt{-{}-bg-from-sources}    & on (rec.)       & Composite onto the real tray inpainted from sources; closes the domain gap \cite{tobin2017domainrand}. \\
\texttt{-{}-neg-frac}           & \texttt{0.10}   & Fraction of background-only scenes to suppress false positives. \\
\texttt{-{}-shadow-dir-mode}    & \texttt{scene}  & One light direction per scene for soft, grounded shadows; \texttt{-{}-no-shadow} disables. \\
\texttt{-{}-shadow-opacity}     & \texttt{0.25}   & Strength of the drop shadow under each seed. \\
\texttt{-{}-edge-feather}       & \texttt{0.5}    & Anti-aliases the paste edge only; never moves labels. \\
\texttt{-{}-degrade}            & on (mild)       & Adds sensor noise, JPEG artefacts, and slight defocus so composites read as photographed \cite{shorten2019augmentation}. \\
\texttt{-{}-noise-sigma}        & \texttt{3}      & Gaussian sensor-noise sigma for the degrade pass. \\
\texttt{-{}-jpeg-quality-min}   & \texttt{80}     & Lower bound of randomized JPEG re-encode quality. \\
\texttt{-{}-jpeg-quality-max}   & \texttt{95}     & Upper bound of randomized JPEG re-encode quality. \\
\texttt{-{}-blur-max-sigma}     & \texttt{0.5}    & Maximum defocus blur applied during degrade. \\
\texttt{-{}-perspective}        & \texttt{0}      & Per-seed perspective warp before placement; \texttt{0.05}--\texttt{0.1} adds viewpoint variety. \\
\texttt{-{}-class-mode}         & \texttt{single} & \texttt{subfolders} maps each top-level source folder to a class; \texttt{single} merges all to one. \\
\texttt{-{}-fmt}                & n/a             & Export format: \texttt{yolo}, \texttt{coco}, \texttt{yolo-seg} (polygons), or \texttt{both} \cite{yolov8,lin2014coco}. \\
\texttt{-{}-masks}              & off             & Also write per-instance masks alongside boxes. \\
\texttt{-{}-num-images}         & n/a             & Number of synthetic scenes to generate (\texttt{0} = verify only). \\
\texttt{-{}-img-size}           & n/a             & Output canvas size in pixels. \\
\texttt{-{}-val-split}          & n/a             & Fraction of generated scenes held out as the validation split. \\
\texttt{-{}-workers}            & \texttt{0} (auto) & Parallel scene generation; output is byte-identical regardless of worker count. \\
\texttt{-{}-parallel-backend}   & \texttt{auto}   & \texttt{loky} processes when assets fit in RAM, else falls back to \texttt{threading} to avoid OOM. \\
\texttt{-{}-seed}               & \texttt{7} (rec.) & Random seed; a fixed seed gives byte-identical output. \\
\texttt{-{}-config}             & n/a             & Load a YAML/JSON baseline; CLI flags still override it. \\
\texttt{-{}-cache-dir}          & \texttt{<out>/.seg\_cache} & Content-hashed segmentation cache shared across runs. \\
\texttt{-{}-qa-report}          & on (rec.)       & Emit QA charts and a visual preview to eyeball before training. \\

\end{longtable}
\end{singlespace}

A caveat carried into this report: the bundled \texttt{seeds/Test} set is
single-seed, so it validates the classifier and the photometric domain but not
multi-seed detection. Synthetic data is for training; a small real, hand-labelled
multi-seed set is kept as the evaluation target.


\chapter{Source Code Excerpt}
\label{appx:code}

The batch state machine is the part of the orchestration that has to stay correct
under concurrency: a multi-image batch fans out into one Celery task per image,
and several workers can run at once. Each batch transition
(\texttt{pending} $\rightarrow$ \texttt{running} and
\texttt{running} $\rightarrow$ terminal) is therefore a compare-and-set update,
not a read-then-write. The transition only commits if the row is still in the
expected state, so exactly one task wins each transition and the others see zero
rows affected and step aside. The excerpt below is a sketch of that update; the
real repository method also stamps \texttt{started\_at} and emits the dual-write
to ClickHouse.

\begin{verbatim}
async def try_transition(session, batch_id, expected, new_status):
    # Compare-and-set: only the first task to arrive flips the state.
    stmt = (
        update(ScanBatch)
        .where(ScanBatch.id == batch_id)
        .where(ScanBatch.status == expected)   # the "compare"
        .values(status=new_status)             # the "set"
        .returning(ScanBatch.id)
    )
    result = await session.execute(stmt)
    won = result.scalar_one_or_none() is not None
    await session.commit()
    return won   # True  -> this task owns the transition
                 # False -> another worker already moved the batch
\end{verbatim}

The first task to flip \texttt{pending} $\rightarrow$ \texttt{running} owns the
start of the batch; the task that records the last image's detection inference
flips \texttt{running} to one of \texttt{succeeded}, \texttt{partial}, or
\texttt{failed}. Because the \texttt{WHERE} clause carries the expected status,
the database serializes the contest and no batch can be advanced twice.
