\chapter{Related Work}

This chapter sets out the background the project builds on. It starts with how
seed quality is checked today, by hand, in a working seed bank. It then surveys
five existing systems that automate part of that job, and notes where each one
falls short of what we need. From there it moves to the computer-vision ideas
behind our pipeline (object detectors and attention classifiers), the
synthetic-data techniques that let us train a detector without labelling
thousands of seeds by hand, and a short review of where AI sits in agriculture
more broadly. The chapter closes with a comparison of prior work against what
Seed Bank delivers.

\section{Traditional methods of seed analysis}

Before a seed batch is accepted into a seed bank for long-term storage, it has
to pass a physical inspection that decides whether the seeds are fit to keep.
For most of the history of seed banking this has been done entirely by hand,
relying on the eye and experience of the person doing the grading.

The standard procedure is a physical examination carried out by a trained
taxonomist or seed-bank technician. It is often called a purity analysis. The
technician looks over the seed sample under a magnifying lens, a
stereomicroscope, or a light table so that small features are easier to see.
The first job is to separate pure seed from inert matter such as soil, chaff,
or plant debris, so that what gets stored is actually seed and not background
material taking up freezer space.

Once the pure seed is isolated, the technician judges each one against a set of
physical criteria: size, shape, surface texture, and colour consistency. Seeds
that look shrivelled, cracked, or discoloured are set aside, as are seeds that
show signs of insect predation or fungal infection. Only seeds that meet the
visual standard for health and structural integrity are passed on to the next
stage. This step matters because a single infected or damaged seed can spoil a
whole storage container, so the cost of missing a bad seed is high.

The weakness of this process is that it is slow, it does not scale, and it
varies between graders. Two technicians can disagree on a borderline seed, and
the same technician can drift over a long shift. That inconsistency, together
with the sheer time it takes, is what motivates an automated alternative.

\section{Existing solutions}

To understand the current landscape we looked at five systems, ranging from
industrial machinery to open-source research code. Each automates some part of
seed inspection, but most of them target growth potential (germination) or
species identification rather than the physical quality checks that long-term
storage actually requires.

\subsection{LemnaTec SeedAIxpert}

The LemnaTec SeedAIxpert \cite{lemnatec2026seedaixpert} is the reference point
for large corporate seed labs and plant breeders. It is an all-in-one hardware
and software product built around an imaging cabinet with controlled lighting,
so that every photograph is taken under identical conditions and the AI software
can produce reproducible scores. It is designed for high throughput, processing
large quantities of seed and replacing subjective human grading with
standardised numbers.

Relevance to our project: the strength of the system is also its barrier, which
is accessibility. It depends on custom, costly hardware and careful calibration
that is impractical for a typical university seed bank or a smaller conservation
effort. Its closed design also means we cannot adapt its algorithms to the
specific physical defects we care about. Our aim is to reach useful accuracy
from ordinary photographs without a dedicated imaging rig.

\subsection{PCS Agri Track}

PCS Agri Track \cite{pcsagri2026track} is a commercial product aimed at
nurseries, built to modernise inventory management and germination tracking.
Unlike the industrial lab equipment it runs on mobile and web, which lowers the
barrier to entry. It moves nurseries away from manual counting by digitising
the tracking process and recording data over time, such as germination curves.

Relevance to our project: the main drawback is its reliance on cloud processing
and a stable internet connection, which is not always available in the field.
Users have also reported that the interface is hard to follow and that the
models generalise poorly, working for common nursery seeds but struggling with
the diverse or rare species a seed bank holds. It still needs manual
intervention for edge cases, which undercuts the goal of automation.

\subsection{Seedy}

Seedy \cite{seedy2026app} is a mobile app for hobbyist gardeners, botanists, and
educators. Its purpose is on-demand seed identification and personal collection
management. A user photographs a seed with a phone camera, and the app's
recognition model returns the species along with metadata and growing guides.

Relevance to our project: Seedy shows that mobile seed analysis is both viable
and easy to use, which supports our own mobile capture flow. Its scope, though,
is strictly identification. It tells the user what a seed is, not how healthy it
is. It does not measure physical metrics or scan for damage and disease, which
is exactly the gap we are trying to fill.

\subsection{GerminationPrediction}

In the academic sphere the GerminationPrediction project of Grimm et al.
\cite{grimm2024germination} is a useful reference. It is open source, uses a
Faster R-CNN detector trained on a public dataset of over twenty-three thousand
images, and automates germination assessment by counting how many seeds in a
batch have sprouted.

Relevance to our project: this is the closest match technically, since it also
uses a CNN-based detector. The biological focus is different, though. They
measure viability (whether a seed will grow), while we measure physical
integrity (whether a seed is sound enough to store). A seed can be viable enough
to sprout and still carry cracks or fungal spores that make it risky to store in
a shared freezer. We borrow their open-source, reproducible approach but retrain
the focus toward defect detection rather than sprout counting.

\subsection{Multi-View Oil Palm system}

The Multi-View Oil Palm project \cite{oilpalmmvcnn} pushes computer vision past
the limits of a single image. It uses Multi-View CNNs (MVCNN) to analyse
irregularly shaped oil palm seeds from several angles, which gives it a 3D
understanding and strong accuracy on that crop.

Relevance to our project: this work makes a point we take seriously, that a
single top-down photograph can miss defects hidden on the side or underside of a
seed. Their multi-view setup gives richer data, but it is specialised for oil
palm and needs a complex rig to capture multiple angles. We aim for a middle
ground: high accuracy from accessible single-view photographs, without the
computational and hardware cost of a multi-view system.

\section{Computer-vision background}

The components below are the model families our pipeline is built from. Seed
quality assessment of a photograph that contains many seeds is really two
problems, not one. First find each seed, then judge it. The first stage is
object detection and the second is per-seed classification. We describe the
ideas behind each here and leave the wiring and results to later chapters.

\subsection{Two-stage detectors}

Two-stage detectors first propose regions that probably contain an object, then
classify and refine each proposal. Faster R-CNN \cite{fasterrcnn2015} is the
reference design. A region proposal network shares features with the detection
head and generates candidate boxes cheaply, which removed the slow external
proposal step that earlier R-CNN variants depended on. Sharing features lets the
whole detector train end to end and run fast enough to be practical.

A backbone produces strong features at a single scale, but seeds in a photograph
vary in size with camera distance and how the kernels are arranged. The Feature
Pyramid Network \cite{lin2017fpn} addresses this by combining the coarse,
semantically strong features from deep layers with the fine, high-resolution
features from shallow layers, giving the detector a consistent representation
across scales at little extra cost. This combination is a sensible default when
detection accuracy matters more than raw speed, which fits a grading workload
where each photo is analysed once and the result is stored.

\subsection{One-stage detectors}

One-stage detectors skip the proposal step and predict boxes and class scores
directly in a single pass. YOLO \cite{redmon2016yolo} introduced the idea by
framing detection as one regression over a grid, trading some localisation
accuracy for a large speed gain. Later versions of the family closed much of
that accuracy gap, and YOLOv8 \cite{yolov8} is a widely used current release
with good tooling for training and export. Seed Bank does not commit to a single
family. The inference layer can run either a fast one-stage detector where
latency matters, for example a conveyor-belt scanning station, or the more
accurate two-stage detector elsewhere, without changing the calling code.

\subsection{Backbones and benchmarks}

Both families rest on a backbone network and are measured against shared public
benchmarks. Residual networks made very deep classifiers trainable by adding
identity skip connections, so gradients reach early layers without vanishing
\cite{resnet2015}. ResNet is the standard backbone choice, and its shallow
member, ResNet-18, is small enough to run many seed crops per image quickly
while still benefiting from pretraining. ImageNet \cite{deng2009imagenet}
provides the large classification dataset that backbones are pretrained on, and
COCO \cite{lin2014coco} provides the detection benchmark with its standard mean
average precision protocol. These datasets matter to us in two concrete ways:
the backbones we use are ImageNet-pretrained, and the synthetic data described
in Section~\ref{sec:ch2-synth} is written in COCO format so generated datasets
drop straight into standard detection training code.

Once a project trains more than one model it also needs a way to track what was
trained, with which data, and how each version scored. MLflow
\cite{mlflow2018} is a common open-source answer for that experiment tracking,
and Seed Bank follows the same pattern when recording and comparing trained
models.

\subsection{Attention in the classifier}

The second stage takes each detected seed, crops it from the original image, and
decides whether that single seed is good or bad. This is binary classification
on a small crop, and our classifier is a ResNet-18 backbone with a Convolutional
Block Attention Module \cite{cbam2018}.

Attention refines what a fixed backbone pays attention to. CBAM is a lightweight
module that learns two attention maps in sequence: a channel map that reweights
which feature channels matter, and a spatial map that reweights which positions
matter. It plugs into an existing backbone with almost no extra parameters and
improved accuracy consistently in the original study. For a seed crop the
attention block helps the classifier focus on the part that carries the verdict,
such as a cracked or discoloured region, rather than spreading its response over
the whole box. The pooled features then feed a small head that produces a
good/bad score. We train a separate classifier per crop type, which is why the
detector groups its boxes by seed type before classification runs.

\section{Synthetic data and cut-and-paste augmentation}
\label{sec:ch2-synth}

This is the area that decided whether the project was feasible. Training the
detector needs images of mixed seeds with a bounding box around every kernel.
Producing that by hand is the labelling bottleneck: an annotator has to click
out hundreds of small, similar, overlapping objects per image, and a useful
training set runs to thousands of images. The companion tool MultiSeedGen
removes that step by generating labelled detection data instead of collecting
it, and its design draws on the synthetic-data literature.

The core technique is cut-and-paste synthesis. Dwibedi et al.
\cite{dwibedi2017cutpaste} showed that pasting cropped object instances onto
varied backgrounds produces training data good enough for competitive
detectors, because a detector mostly needs to see an object's local appearance
in many contexts, not a globally coherent scene. The paste does not have to look
realistic to a human. As long as the local patches are varied, the detector
still generalises to real images. The payoff is that the label comes for free:
when the generator places a seed cut-out at a known position and scale, it
already knows the exact bounding box, so every synthetic image is perfectly
annotated with no human in the loop.

Cut-and-paste only works if the object crops are clean, with the background
removed down to the seed's true outline. For the easy cases, classical
thresholding handles high-contrast backgrounds with no extra dependencies. For
feathered edges the U\textsuperscript{2}-Net salient-object model
\cite{qin2020u2net} separates foreground from background, and for the hardest
cases Segment Anything \cite{kirillov2023sam} produces high-quality masks from a
box or point prompt. The learned backends are optional, so a basic install stays
light, and low-quality masks are dropped before they reach the compositing step.

A detector trained on synthetic images still has to transfer to real
photographs. The standard answer is domain randomization: vary the synthetic
data so widely that the real distribution looks like just another sample. Tobin
et al. \cite{tobin2017domainrand} introduced the idea for simulation, randomising
textures, lighting, and placement so a model trained only on synthetic data
transfers to the real world. MultiSeedGen applies the same idea by randomising
backgrounds, per-seed geometry (scale, rotation, flip, shear, perspective), and
photometric conditions (gamma, blur, noise, vignetting) so the generated set
spans far more variation than a hand-collected set would. These transforms sit
inside the broader augmentation literature surveyed by Shorten and Khoshgoftaar
\cite{shorten2019augmentation}, which finds that geometric and photometric
perturbation reduces overfitting across vision tasks, and libraries such as
Albumentations \cite{buslaev2020albumentations} made these transforms fast and
annotation-aware, which matters because moving an image must move its boxes with
it.

One design point is worth calling out because it is easy to get wrong. The
generator enforces a deterministic train/validation split in which a source seed
reserved for validation never appears in any training image. This is the
leakage-safe split. Without it the model in effect sees the validation seeds
during training and reports an inflated score, so the split has to be decided at
the source-seed level rather than the image level.

\section{AI in agriculture}

AI is now a routine part of modern agriculture rather than an experiment. The
shift is driven by pressure on the sector: arable land is shrinking and climate
conditions are more volatile, so the older "yield at all costs" mindset has given
way to a precision-first one, where efficiency and resilience matter as much as
raw output. The survey by Kamilaris and Prenafeta-Bold\'u
\cite{kamilaris2018deepag} documents this move across agriculture, from yield
prediction to weed and crop classification to quality grading, all shifting from
hand-crafted features to learned representations with steady accuracy gains where
labelled data was available.

In precision farming, models combine satellite imagery, soil-sensor readings,
and local weather forecasts into field-level prescriptions for how much water or
input each area needs. Closer to our problem is fine-grained computer vision for
quality control. Where ordinary detection just identifies a crop type,
fine-grained methods pick up subtle within-class differences such as hairline
fractures, surface discoloration, or early-stage fungal damage that often escape
the human eye. Mohanty et al. \cite{mohanty2016plant} show both the promise and
the catch: their CNNs reached high accuracy at recognising crop diseases from
leaf photographs on a curated dataset, but accuracy dropped sharply on images
taken under conditions different from the training set.

That last result points to the barriers the field still faces. The first is
capital cost: sensors and autonomous machinery carry a high upfront price that
keeps smaller operations out. The second is the digital divide, a widening gap
between fully integrated commercial farms and smallholders. The third is the one
that shaped our own work most: dataset variability and domain generalization.
Seed diversity and lighting changes break models trained on a narrow set, and
the usual fix is to collect local data and retrain for the specific harvest. The
gap between a clean benchmark and field photographs is the central practical
problem for any agricultural vision system, and it is the reason we treat data
variety as a first-class concern rather than an afterthought.

\section{Comparative analysis}

The components above are individually well established. What distinguishes this
project is not a new model but how the pieces are assembled into one workable
pipeline, and the decision to remove the data-labelling bottleneck with a
purpose-built generator rather than working around it.

Most prior work in seed and grain vision is a single artifact: one classifier or
one detector, trained against a hand-labelled dataset and evaluated once. Such a
system answers one question, either "where are the seeds" or "is this seed good",
but not the combined question the domain actually asks, which is how many seeds
are in this photo and how many of them are bad. It is usually trained on a
hand-labelled dataset, which caps the dataset size and therefore the accuracy.

Seed Bank differs on each point. It chains a detector and a per-type classifier
into one detect-then-classify pipeline, so a single photograph yields a located,
per-seed good/bad verdict and an aggregate report. It supports multiple crop
types, each with its own classifier, with detections grouped by type before the
classification stage. And the labelling bottleneck is attacked at the source by
MultiSeedGen, which generates perfectly labelled multi-seed data with controlled
variety, so the detector can be trained at a scale that hand-labelling could not
reach. Table~\ref{tab:ch2-compare} summarises the contrast.

\begin{singlespace}
\begin{longtable}{p{0.30\textwidth} p{0.31\textwidth} p{0.31\textwidth}}
\caption{Typical prior work in seed and grain vision versus the Seed Bank platform.}
\label{tab:ch2-compare} \\
\toprule
\textbf{Dimension} & \textbf{Typical prior work} & \textbf{Seed Bank} \\
\midrule
\endfirsthead
\toprule
\textbf{Dimension} & \textbf{Typical prior work} & \textbf{Seed Bank} \\
\midrule
\endhead
\bottomrule
\endfoot
Scope &
Single model: a detector \emph{or} a classifier &
Two-stage detect-then-classify pipeline producing per-seed verdicts and an aggregate report \\
Quality question answered &
``where are the seeds'' \emph{or} ``is this seed good'' &
``how many seeds, and how many are bad'' over a mixed photo \\
Crop types &
Usually one species, fixed at training &
Multiple crop types, each with its own classifier; detections grouped by type \\
Training data &
Hand-labelled, which caps dataset size &
Synthetic cut-and-paste generation with free, exact labels and controlled variety \\
Data integrity &
Leakage and label errors are easy to miss &
Leakage-safe split at the source-seed level, with controlled overlap and box clipping \\
\end{longtable}
\end{singlespace}

The wider point is that the contribution sits at the system level. Each model
family here is taken from published work and used as intended. The originality is
in binding detection, per-type classification, and a synthetic-data generator
into one tool that a non-researcher can run, and in attacking the labelling
bottleneck head-on rather than living with a small hand-labelled dataset. The
chapters that follow describe how that system is built and how it performs; the
experimental numbers are reported in the evaluation in Chapter~\ref{ch:testing}.
